---
title: "Configuration Options for Parsing from JSON"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Configuration Options for Parsing from JSON}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>"
)
```

```{r setup}
suppressPackageStartupMessages({
  library(yyjsonr)
})
```

Overview
-----------------------------------------------------------------------------

This vignette:

* introduces the `opts` argument for reading JSON with the `read_json_X()` family 
  of functions.
* outlines the creation of default options with `opts_read_json()`
* provides extended examples of how these options control parsing of JSON


The `opts` argument - Specifying options when reading JSON
-----------------------------------------------------------------------------

All `read_json_x()` functions have an `opts` argument.  `opts` takes 
a named list of options used to configure the way `yyjsonr` parses JSON 
into R objects.

The default argument for `opts` is an empty list, which internally sets the 
default options for parsing.

The default options for parsing can also be viewed by running
`opts_read_json()`.

The following three function calls are all equivalent ways of calling
`read_json_str()` using the default options:

```{r eval=FALSE}
read_json_str(str)
read_json_str(str, opts = list())
read_json_str(str, opts = opts_read_json())
```


Setting arguments to override the default options
-----------------------------------------------------------------------------

Setting a single option (and keeping all other options at their default value)
can be done in a number of ways.

The following three function calls are all equivalent:

```{r eval=FALSE}
read_json_str(str, opts = list(str_specials = 'string'))
read_json_str(str, opts = opts_read_json(str_specials = 'string'))
read_json_str(str, str_specials = 'string')
```




Option `promote_num_to_string` - mixtures of numeric and string types
-----------------------------------------------------------------------------

By default, `yyjsonr` does not promote string values to numerica values i.e.
`promote_num_to_string = FALSE`.

If an array contains mixed types, then an R *list* will be returned, so that
all JSON values retain their original type.

```{r}
json <- '[1,2,3.1,"apple", null]'
read_json_str(json)
```

If `promote_num_to_string` is set to `TRUE`, then `yyjsonr` will promote 
numeric types to strings if the following conditions are met:

* values are stored in a JSON array
* the JSON array only contains numerics, strings or the JSON `null` value


```{r}
yyjsonr::read_json_str(json, promote_num_to_string = TRUE)
```



Option `df_missing_list_elem` - Missing list elements (when parsing data.frames)
-----------------------------------------------------------------------------

When JSON data is being parsed into an R data.frame some columns become
*list-columns* if there are mixed types in the original JSON.

It is possible that some values are completely missing in the JSON 
representation, and the `df_missing_list_elem` specifies the replacement for 
this missing value in the R data.frame.  The default value is 
`df_missing_list_elem = NULL`.

### JSON to data.frame (no *list columns* needed)

```{r}
str <- '[{"a":1, "b":2}, {"a":3, "b":4}]'
read_json_str(str)
```


### JSON to data.frame - *list-columns* required

```{r}
str <- '[{"a":1, "b":[1,2]}, {"a":3, "b":2}]'
read_json_str(str)
```


```{r}
str <- '[{"a":1, "b":[1,2]}, {"a":2}]'
read_json_str(str)
read_json_str(str, df_missing_list_elem = NA)
```







Option `obj_of_arrs_to_df` - Reading JSON as a data.frame
-----------------------------------------------------------------------------

By default, if JSON looks like it represents a data.frame it will be loaded
as such. That is, a JSON `{}` object which contains only `[]` arrays (all of
equal length) will be treated as data.frame.  This is the default i.e.
`obj_of_arrs_to_df = TRUE`.

If `obj_of_arrs_to_df = FALSE` then this data will be read in as a named list.
In addition, if the `[]` arrays are not all the same length, then the data 
will also be read in as a named list as no inference
of missing values will be done.

```{r}
str <- '{"a":[1,2],"b":["apple", "banana"]}'
read_json_str(str)
read_json_str(str, obj_of_arrs_to_df = FALSE)
```


```{r}
str_unequal <- '{"a":[1,2],"b":["apple", "banana", "carrot"]}'
read_json_str(str_unequal)
```


Option `arr_of_objs_to_df` - Reading JSON as a data.frame
-----------------------------------------------------------------------------

```{r}
str <- '[{"a":1, "b":2}, {"a":3, "b":4}]'
read_json_str(str)
read_json_str(str, arr_of_objs_to_df = FALSE)
```

```{r}
str <- '[{"a":1, "b":2}, {"a":3, "b":4, "c":99}]'
read_json_str(str)
```


Option `str_specials` - Reading string `"NA"` from JSON
-----------------------------------------------------------------------------

JSON only really has the value `null` for representing special missing values, 
and this is converted to an R `NA_character_` value when it is encountered in 
a string-ish context.

When `yyjsonr` encounters a literal `"NA"` value in a string-ish context, 
its conversion to an R value is controlled by the `str_specials` options


The possible values for the `str_specials` argument are:

* `string` read in as the literal character string `"NA"`  (the default behaviour)
* `special` read in as `NA_character_`

```{r}
str <- '["hello", "NA", null]'
read_json_str(str) # default: str_specials = 'string'
read_json_str(str, str_specials = 'special')
```


Option `num_specials` - Reading numeric `"NA"`, `"NaN"` and `"Inf"`
-----------------------------------------------------------------------------

JSON only really has the value `null` for representing special missing values, 
and this is converted to an R `NA_integer_` or `NA_real_` value when it is encountered in 
a number-ish context.

When `yyjsonr` encounters a literal `"NA"`, `"NaN"` or `"Inf"` value in a number-ish context, 
its conversion to an R value is controlled by the `num_specials` options.


The possible values for the `num_specials` argument are:

* `special` read in as an actual numeric `NA`, `NaN` or `Inf` value (the default behaviour)
* `string` read in as the literal character string `"NA"` etc

```{r}
str <- '[1.23, "NA", "NaN", "Inf", "-Inf", null]'
read_json_str(str) # default: num_specials = 'special'
read_json_str(str, num_specials = 'string')
```



Option `int64` - large integer support
-----------------------------------------------------------------------------

JSON supports large integers outside the range of R's 32-bit integer type.

When such a large value is encountered in JSON, the `int64` option controls the
value's representation in R.

The possible values for the `int64` option are:

* `string` store JSON integer as a string in R
* `double` will store the JSON integer as a double precisision numeric. If the 
  integer is outside the range +/- 2^53, then it may not be stored perfectly in
  the double.
* `bit64` convert to a 64-bit integer supported by the [`{bit64}`](https://cran.r-project.org/package=bit64)
  package.

```{r echo=FALSE}
suppressPackageStartupMessages(
  library(bit64)
)
```


```{r}
str <- '[1, 274877906944]'

# default: int64 = 'string'
# Since result is a mix of types, a list is returned
read_json_str(str) 

# Read large integer as double
robj <- read_json_str(str, int64 = 'double')
class(robj)
robj

# Read large integer as 'bit64::integer64' type
library(bit64)
read_json_str(str, int64 = 'bit64')
```




Option `length1_array_asis` - distinguishing scalars from length-1 vectors
-----------------------------------------------------------------------------

JSON supports the concept of both scalar and vector values i.e. in JSON 
scalar `67` is different from an array of length 1 `[67]`.  The 
`length1_array_asis` option is for situations where it is important to 
distinguish these value types in R.

However, R does not make this distinction between scalars and vectors of length 1.

To assist in translating objects from JSON to R and back to JSON, setting
`length1_array_asis = TRUE` will mark JSON arrays of length 1 with the 
class `AsIs`.  This option defaults to `FALSE`.

```{r}
read_json_str('67')   |> str()
read_json_str('[67]') |> str()

read_json_str('67'  , length1_array_asis = TRUE) |> str()
read_json_str('[67]', length1_array_asis = TRUE) |> str() # Has 'AsIs' class
```


This option is then used with the option `auto_unbox` when writing JSON in 
order to control how length-1 R vectors are written.  Shown below, if the
length-1 vector is marked with `AsIs` class when reading, then when writing
out to JSON with `auto_unbox = TRUE` it becomes a JSON vector value.


In the following example, only the second value (`[67]`) is affected by the option
`length1_array_asis`.  When the option is `TRUE` the value is tagged with a 
class of `AsIs`.  Then when the created R object is subsequently written out to a JSON string,
its structure is determined by `auto_unbox` which understands how to handle
this class.

```{r}
str <- '{"a":67, "b":[67], "c":[1,2]}'

# Length-1 vectors output as JSON arrays
read_json_str(str) |>
  write_json_str(auto_unbox = FALSE) |>
  cat()

# Length-1 vectors output as JSON scalars
read_json_str(str) |>
  write_json_str(auto_unbox = TRUE) |>
  cat()

# Length-1 vectors output as JSON arrays
read_json_str(str, length1_array_asis = TRUE) |>
  write_json_str(auto_unbox = FALSE) |>
  cat()

# !!!!
# Those values marked with 'AsIs' class when reading are output
# as length-1 JSON arrays
read_json_str(str, length1_array_asis = TRUE) |>
  write_json_str(auto_unbox = TRUE) |>
  cat()
```





Option `yyjson_read_flag` - internal `YYJSON` C library options
-----------------------------------------------------------------------------

The `yyjson` C library supports a number of internal options for reading JSON.

These options are considered advanced, and the user should read the
original `yyjson` documentation for further explanation on what they control.

**Warning**: some of these advanced options do not make sense for interfacing with R,
or otherwise conflict with how this package converts JSON to R objects. 

```{r}
# A reference list of all the possible YYJSON options
yyjsonr::yyjson_read_flag

read_json_str(
  "[1, 2, 3, ] // A JSON comment not allowed by the standard",
  opts = opts_read_json(yyjson_read_flag = c(
    yyjson_read_flag$YYJSON_READ_ALLOW_TRAILING_COMMAS,
    yyjson_read_flag$YYJSON_READ_ALLOW_COMMENTS
  ))
)
```
