---
title: "Comparison to jsonlite parsing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparison to jsonlite parsing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(yyjsonr)
```



Parsing differences compared to `{jsonlite}`
=============================================================================

`{jsonlite}` and `{yyjsonr}` may read and write some JSON differently
due to varying assumptions, data configurations or option settings.

This document keeps a record of major differences to be aware of.



In `yyjsonr` 3-d arrays are parsed as multiple 2-d matrices and combined
-----------------------------------------------------------------------------

In `{yyjsonr}` the order in which elements in an array are serialized to 
JSON correspond to a JSON `[]` array of row-major matrices in human-readable order.

`{jsonlite}` does things differently. 

The array formats are internally
consistent within each package, but not cross-compatible between them i.e.
you cannot serialize an array in `{yyjsonr}` and re-create it exactly 
using `{jsonlite}`.

In the examples below, a simple 3d matrix is serialized with both 
`jsonlite` and `yyjsonr`.

```{r}
# A simple 3D array 
mat <- array(1:12, dim = c(2,3,2))
mat
```


```{r}
# jsonlite's serialization of matrices is internally consistent and re-parses
# to the initial matrix.
str <- jsonlite::toJSON(mat, pretty = TRUE)
cat(str)
jsonlite::fromJSON(str)
```


```{r}
# yyjsonr's serialization of matrices is internally consistent and re-parses
# to the initial matrix.
# But note that it is *different* to what jsonlite does.
str <- yyjsonr::write_json_str(mat, pretty = TRUE)
cat(str)
yyjsonr::read_json_str(str)
```
